---
title: "Coursera - Practical Machine Learning - Course Project"
output: html_document
---
### Summary
Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

*The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

#### Methodology
I will build a random forest, decision tree, and Gradient boosted model to predict the out of sample data. I will select the best performing model.

Load Required Libraries
```{r}
library(knitr)
library(caret)
library(rpart)
library(randomForest)
set.seed(12345)
```

Import Data
```{r}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))
```

Split Data for modeling
```{r}
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
```

Clean Data
```{r}
# remove Near Zero Varianace Variables
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]

# remove N/A Variables
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]

# remove ID Variables
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
```

Random Forest
```{r}
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf", trControl=controlRF)
modFitRandForest$finalModel
```

```{r}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

Decision Tree
```{r}
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
```

```{r}
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

GBM
```{r}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm", trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```

```{r}
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM
```

Model Selection
```{r}
confMatRandForest
confMatDecTree
confMatGBM
```

#### Result
The accuracy of the models was as follows:

The Random Forest Model was 98.54% accurate
The Decision Tree Model was 65.59% accurate
The Gradient Boosted Model was 92.56% accurate

Based on these results I will select the Random Forest Model.

Predict out of sample data.
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```